---
title: 'ESM242 MiniGP: Fisheries -- Model Component'
author: "Emma Tao, Ryan Anderson, Jackson Hayes"
date: "2024-11-17"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Model Specification

## Objective

We are going to model the fishing efforts of a recreational fishing company
on two hypothetical reefs. The objective is going to be to maximize the total
number of fish caught throughout the entire timespan simulated.

We will simulate 50 years, which is about the lifespan of a boat.

The constraint is that there will be a finite amount of fishing trips that can
occur in any given month.

This problem was inspired by 2 real reefs of approximate equal size (1 acre) in the Santa Barbara channel.

## Stock Modeling

The basic stock equation is 

$$ X_{t+1} = X_t + F(X_t) - H_t + M(X_t) $$
- $F(X_t)$ is the amount of fish added to the stock at period t due to reproductive growth
- $H_t$ is fish harvested in period $t$
- $M(X_t)$ is the number of fish added to the stock in period t due to migration

We'll model $F(X_t)$ (or the amount of fish added to the stock at period t) with the equation

$$ F(X_t) = rX_t(1-\frac{X_t}{K})$$

We'll model $M(X_t)$ (or the number of fish added to the stock in period t due to migration) with the equation

$$ M(X_t) = z(K-X_t) $$

Where $z$ is some set parameter.


## Harvest Modeling

The amount of harvest that occurs in a given period will be the sum of the harvests that occur at 
the two individual reefs.

$$ H_t = q(E_{1t}X_{1t} + E_{2t}X_{2t}) $$
Where q is some set parameter.

We'll have a constraint on the total fishing effort that can go across the 2
reefs:

$$ E_{1t} + E_{2t} \le E_{max} $$
Where E is some constant.


## Objective Function

Our objective is going to be to maximize

$$ ln \sum_{t=1}^T H_t $$

 $$ \sum_{t=1}^T \rho^t ln(aH_t) $$

by choosing optimal $E_{1t}$ and $E_{2t}$ values.

The constant $a$ is a parameter chosen to scale values of H_t such that the differences will be large enough to be used by the optimizer.

$\rho$ is the discount term.




## Model Parameters
| **Name**                | **Symbol** | **Estimate**        | **Description**                                      |
|-------------------------|------------|---------------------|------------------------------------------------------|
| Initial Stock           | $X_0$      | ~1000 fish          | Same starting value for both reefs.                 |
| Carrying Capacity       | $K$        | ~2000 fish          | Max population the environment supports.            |
| Growth Rate             | $r$        | 0.1-0.5 $yr^{-1}$   | Intrinsic growth rate for whitefish populations.     |
| Fishing Effort Cap      | $E_{max}$  | ~15 fishing days    | Max monthly fishing days.                           |
| Harvest Constant        | $q$        | ~0.05             | Effectiveness of fishing effort; currently calculated assuming 50 fish caught per trip, with 15 trips per week (when stock is 1000 fish) |
| Migration Constant      | $z$        | 0.2                 | Proportion of remaining carrying capacity that becomes occupied by immigrating fish in the next period.         |
| Simulation Time         | $T$        | 24                 | Periods (months) to simulate.                       |
| Discount Term           | $\rho$     | 0.9                 | Discount term for objective function.                |
| Utility Scaling Constant| $a$        | 1                   | Scaling constant for objective function.             |


# Model Implementation



## Setting Up

```{r}
# Clear Env
rm(list = ls())

# Load Libraries
library(tidyverse)
library(nloptr)
library(knitr)

# Set seed for reproducibility
set.seed(1234)
```

```{r}
# Load the model functions
source("fishery_model.R")
source("optimization.R")
source("model_viz.R")

# x0, max_effort, initial_stock, carrying_capacity, 
                                   # r, q, z, num_periods, 
                                   # rho, util_scaling_constant
# Now let's set up the parameters for the model
x0 <- 5
max_effort <- 15
initial_stock <- 1000
carrying_capacity <- 2000
r <- 0.1
q <- 0.05
z <- 0.2
num_periods <- 24
rho <- 0.95
util_scaling_constant <- 1

# Run the optimization
optimal_values <- optimize_fishery_model(x0, max_effort, initial_stock, carrying_capacity, r, q, z, num_periods, rho, util_scaling_constant)

# visualize results


```



## Inspecting Results

Let's see what's going on with the optimal values from the model run. Is it what we expect

```{r}
# Run the model with opimal values
optimal_run <- run_model(optimal_values[1:SIMULATION_LENGTH], optimal_values[(SIMULATION_LENGTH+1):(2*SIMULATION_LENGTH)], EFFORT_CAP, INITIAL_STOCK, CARRYING_CAPACITY, GROWTH_RATE, HARVEST_CONSTANT, MIGRATION_CONSTANT, SIMULATION_LENGTH, RHO, UTIL_SCALING_CONSTANT)
# optimal_run <- run_model(, EFFORT_CAP, INITIAL_STOCK, CARRYING_CAPACITY, GROWTH_RATE, HARVEST_CONSTANT, MIGRATION_CONSTANT, SIMULATION_LENGTH)

# Plot the results

plot(optimal_run$stock1, type = "l", xlab = "Time", ylab = "Stock", main = "Stock 1 Over Time")
plot(optimal_run$stock2, type = "l", xlab = "Time", ylab = "Stock", main = "Stock 2 Over Time")

# plot(optimal_run$harvest_1, type = "l", xlab = "Time", ylab = "Harvest", main = "Harvest 1 Over Time")
# plot(optimal_run$harvest_2, type = "l", xlab = "Time", ylab = "Harvest", main = "Harvest 2 Over Time")

plot(optimal_values[1:SIMULATION_LENGTH], type = "l", xlab = "Time", ylab = "Effort", main = "Effort 1 Over Time")
plot(optimal_values[(SIMULATION_LENGTH+1):(2*SIMULATION_LENGTH)], type = "l", xlab = "Time", ylab = "Effort", main = "Effort 2 Over Time")

# Plot harvest 1 and harvest 2 over time together on one graph
# use ggplot
harvest_data <- data.frame(time = 1:SIMULATION_LENGTH, harvest1 = optimal_run$harvest_1, harvest2 = optimal_run$harvest_2)
harvest_data <- gather(harvest_data, key = "reef", value = "harvest", -time)

ggplot(harvest_data, aes(x = time, y = harvest, color = reef)) + geom_line() + labs(title = "Harvest Over Time", x = "Time", y = "Harvest")


```


